---
title: "Physical Fractions"
author: "Kaelin Sylva"
layout: post
output:
  html_document:
    df_print: paged
tags: ["aggregate size"],["partical size"],["fractions"]
category: Secondary controls on carbon turnover
level 1:"4"
level 2:"a"
level 3: "ii"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Quantitative Methods - Physical Fractions

Quantitative methods can be used to further study soil organic matter and its behavior. More importantly, utilizing fractionation processes can aid our understanding of the turnover process of soil organic matter. When conducting studies or projects, there are various techniques that can be utilized to fractionate soils. Fractionation is the physical separation of different components within the soil, which can be done physically, biologically, or chemically. Within this section, physical fractions will be discussed. For more information on both the chemical and biological process, please refer to the following links: , respectively. Through physically fractionation, soils will be sorted by their size and potentially by their density (Gregorich et al, 2006). Utilizing one type of fractionation can answer a few characteristic of soil organic matter; however, utilizing multiple fractions can provide a more in depth understanding of that particular sample. More in depth information regarding some of the tools and equipment utilized can be found at here . 

```{r infobox1, include=FALSE}
#I've decided to use native R to discuss the syntax and formatting of this example template.  By using include=FALSE in the R chunk definition I have effectively kept it from being included in the final product.  When using this template the template intoduction and all R chunks with "infobox" labels can be removed. 
```
#The Conundrum of SOM

Outline the history of how SOC/SOM has been conceptualized

List of important publications:

Sollins et al. (1996) DOI: [10.1016/S0016-7061(96)00036-5](https://doi.org/10.1016/S0016-7061(96)00036-5)

#Humification as a theory is unsupported

Explain how humication theory is unsupported.

![Figure 1. How our view of SOM has changed as we employ new soil measurement techniques. For full figure text see Schmidt et al. (2011) DOI: [10.1038/nature10386](https://www.nature.com/articles/nature10386)](https://media.springernature.com/m685/nature-assets/nature/journal/v478/n7367/images/nature10386-f2.2.jpg)

```{r infobox2, include=FALSE}
#above is the first example of inserting an image into your Rmarkdown file.  the '![]()' syntax is used.  You start with an '!'.  Then Within the brakets [] you include the text that you want included as the figure caption.  Then Within the parentheses () you include the link to where the image can be found; the image is from the web in the above example.  I have also nested the link to Schmidts paper's DOI within the figure text which is why it appears complicated.  Below is a similar example with a different image.  Remember you will be copying this template and, though trial and error, it will begin to make sense... and if it doesnt, we'll help you in class. 
```

Discuss the emerging understandings in stabilization mechanisms. 

![Figure 2. An emerging understanding of stabilization/destabilization processes has replaced humification theory. For full figure text see Schmidt et al. (2011) DOI: [10.1038/nature10386](https://www.nature.com/articles/nature10386)](https://media.springernature.com/m685/nature-assets/nature/journal/v478/n7367/images/nature10386-f3.2.jpg)

List of important publications:

Schmidt et al. (2011) DOI: [10.1038/nature10386](https://www.nature.com/articles/nature10386)

#Rethinking the Conceptual Theory of SOM Sequestration

As paradigms have shifted, its important to get away from humification terminology and embrace mechanistic theories and conceptualizations of soil carbon stabilization.

![Figure 3. The soil continum model (SCM) is an attempt to bring together disparate conceptualizations of soil carbon storage, which is an important step forward in soil science as we try to determine the best measurement and modeling techniques to predict SOC changes. For full figure text see Lehmann & Kleber (2015) DOI: [10.1038/nature16069](https://www.nature.com/articles/nature16069)](https://media.springernature.com/m685/nature-assets/nature/journal/v528/n7580/images/nature16069-f2.jpg)

List of important publications:

Lehmann & Kleber (2015) DOI: [10.1038/nature16069](https://www.nature.com/articles/nature16069)

#Moving forward: bringing together theory, measurement, and modeling

Theory, measurement, and modeling need to be thought of as a single process to close the loop of the scientific method and create usable products to not just inform further science, but also provide farmers, land managers, policy makers, and the general public a tool to understand soil carbon. 

List of important publications:

Triangle paper
