---
title: "Thermal Fractionation"
author: "Kristy Lam"
layout: post
output: word_document
tags: ["thermal fractionation"]
category: "Secondary controls on carbon turnover"
level1: "Quantitative Methods"
level2: "Fractionation of SOM"
level3: "Thermal Fractionation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r infobox1, include=FALSE}
#I've decided to use native R to discuss the syntax and formatting of this example template.  By using include=FALSE in the R chunk definition I have effectively kept it from being included in the final product.  When using this template the template intoduction and all R chunks with "infobox" labels can be removed. 
```
#Thermal Fractionation

Another technique that is increasingly being used by soil scientists is thermal or spectral fractionation. In general chemistry, we learned that all matter emits electromagnetic radiation. The electromagnetic spectrum shows the scale of wavelengths are emitted (see image below).



![Figure 1. Electromagnetic Spectrum](TG&DSC(2).png) (https://en.wikipedia.org/wiki/File:EM_Spectrum_Properties_edit.svg#filelinks)

Thermal or spectral fractionation utilizes heat (thermal) or a particular type of electromagnetic radiation (spectral) to separate soil organic matter (SOM), then identify what types of soil organic carbon (SOC) are present in a given soil sample. However, these methods often require adequate statistical analyses techniques (i.e. Partial Least Squares Regression) that can sort through many highly collinear spectral bands from relatively few observations (Denef, Plante, and Six, 2009). Recent studies (see "Applications" section) have utilized these techniques to their advantage to answer relevant questions about SOM dynamics and stability.

#Why We Use These Methods

SOM includes a variety of different forms of carbon that are produced from heterogeneous sources (i.e. microbes processes, weathering, etc). There is no one method that can measure and identify every single carbon component in soil. The advantage of thermal or spectral fractionation is that these techniques combined can measure and identify SOM/SOC. Furthermore, these methods are often coupled with another (i.e. biological, chemical, physical fractionation - include link to these pages) to produce results that have important implications toward soil health and soil best management practices. Recent studies are starting to delve deep into the thermal and spectral fraction to answer questions about soil carbon dynamics and SOM stability. Although there are many thermal/spectral techniques out there, we will focus on a few that are often used when studying SOM/SOC.



***

>“...thermal analysis is a highly integrative method that requires little sample preparation and allows the analysis of bulk soil samples without pretreatment, even with low C content. Using thermal analysis techniques to investigate SOM stability is highly attractive.”   
>Peltre et al., (2013) DOI: [10.2136/sssaj2013.02.0081](https://www.soils.org/publications/sssaj/abstracts/77/6/2020)


***

>“Analysis of the results revealed the suitability of thermal methods to provide simple, quick and reliable information on organic matter stabilisation during the composting process of agroindustrial residua.”    
>Dell’Abate et al., (2000) DOI: [10.1023/A:1010157115211](https://link.springer.com/article/10.1023/A%3A1010157115211)

***



#Thermal Fractionation Methods

##A. Thermogravimetry (TG) & Differential Scanning Calorimetry (DSC)

This method involves two parts: thermogravimetry (TG) and differential scanning calorimetry (DSC). The former involves increasingly heating a sample in an inert or oxidative gas atmosphere that result in weight loss or gain. The changes in sample weight can be attributed to alterations in chemical or physical properties, thus TG is important in investigating thermal stability of SOM. The latter will plot a figure according to heat flow (i.e. exotherm vs endotherm reaction). DSC analyses do not require large samples, numerous controls, or extensive method development. In addition, this technique can be applied to whole soil sampling without requirement of pretreatments (i.e. chemical extraction) or isolated fractions.Check out this [video from Setaram](http://us.setaram.com/en/setaram-products/couplings-gas-analysis-2/tga-sta-ms/sensys-evo-tg-dsc-ega/1/).

Sources:   
EAG Laboratory (https://www.eag.com/)



#Applications

**SOM stability**

In 2013, Peltre et al. further assessed the relationships between thermal and biological indices of SOM stability using a wide range of soils with varying C concentrations collected from across the United States. Soils were analyzed using TG-DSC coupled with CO2–EGA. These measurements were compared with SOM biological stability measured during long-term laboratory incubations. Results showed that different relationships exist between the thermal and biological stabilities of SOM (i.e. differences in C concentrations), which can be attributed to different mechanisms of SOM stabilization. Thus, mineral association of SOM and  abundance of intact plant debris are vital in low-C soils and high-C soils, respectively.
DOI: [10.2136/sssaj2013.02.0081](https://www.soils.org/publications/sssaj/abstracts/77/6/2020)

```{r infobox2, include=FALSE}
#above is the first example of inserting an image into your Rmarkdown file.  the '![]()' syntax is used.  You start with an '!'.  Then Within the brakets [] you include the text that you want included as the figure caption.  Then Within the parentheses () you include the link to where the image can be found; the image is from the web in the above example.  I have also nested the link to Schmidts paper's DOI within the figure text which is why it appears complicated.  Below is a similar example with a different image.  Remember you will be copying this template and, though trial and error, it will begin to make sense... and if it doesnt, we'll help you in class. 
```

