---
title: "Thermal and Spectral Fractionation"
author: "Kristy Lam"
layout: post
output: word_document
tags: ["thermal fractionation", "spectral fractionation",]
category: "Secondary controls on carbon turnover"
level1: "Quantitative Methods"
level2: "Quantitative Methods"
level3: "Spectral or Thermal Fractionation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r infobox1, include=FALSE}
#I've decided to use native R to discuss the syntax and formatting of this example template.  By using include=FALSE in the R chunk definition I have effectively kept it from being included in the final product.  When using this template the template intoduction and all R chunks with "infobox" labels can be removed. 
```
#Thermal or Spectral Fractionation

Another technique that is increasingly being used by soil scientists is thermal or spectral fractionation. In general chemistry, we learned that all matter emits electromagnetic radiation. The electromagnetic spectrum shows the scale of wavelengths are emitted (see image below).



![Figure 1. Electromagnetic Spectrum](TG&DSC(2).png) (https://en.wikipedia.org/wiki/File:EM_Spectrum_Properties_edit.svg#filelinks)

In the context of soil science, thermal or spectral fractionation utilizes heat (thermal) or a particular type of electromagnetic radiation (spectral) to separate soil organic matter (SOM), then identify what types of soil organic carbon (SOC) are present in a given soil sample. However, these methods often require adequate statistical analyses techniques (i.e. Partial Least Squares Regression) that can sort through many highly collinear spectral bands from relatively few observations (Denef, Plante, and Six, 2009). Recent studies (see "Applications" section) have utilized these techniques to their advantage to answer relevant questions about SOM dynamics and stability.

#Why We Use These Methods

SOM includes a variety of different forms of carbon that are produced from heterogeneous sources (i.e. microbes processes, weathering, etc). There is no one method that can measure and identify every single carbon component in soil. The advantage of thermal or spectral fractionation is that these techniques combined can measure and identify SOM/SOC. Furthermore, these methods are often coupled with another (i.e. biological, chemical, physical fractionation - include link to these pages) to produce results that have important implications toward soil health and soil best management practices. Recent studies are starting to delve deep into the thermal and spectral fraction to answer questions about soil carbon dynamics and SOM stability. Although there are many thermal/spectral techniques out there, we will focus on a few that are often used when studying SOM/SOC.

“...thermal analysis is a highly integrative method that requires little sample preparation and allows the analysis of bulk soil samples without pretreatment, even with low C content. Using thermal analysis techniques to investigate SOM stability is highly attractive.” (Peltre et al., 2013) DOI: [10.2136/sssaj2013.02.0081]

“Analysis of the results revealed the suitability of thermal methods to provide simple, quick and reliable information on organic matter stabilisation during the composting process of agroindustrial residua.” (Dell’Abate et al., 2000) DOI: [10.1023/A:1010157115211] 




![Figure 1. How our view of SOM has changed as we employ new soil measurement techniques. For full figure text see Schmidt et al. (2011) DOI: [10.1038/nature10386](https://www.nature.com/articles/nature10386)](https://media.springernature.com/m685/nature-assets/nature/journal/v478/n7367/images/nature10386-f2.2.jpg)




#Thermal/Spectroscopic Methods

##A. Infrared (IR) Spectroscopy



```{r infobox2, include=FALSE}
#above is the first example of inserting an image into your Rmarkdown file.  the '![]()' syntax is used.  You start with an '!'.  Then Within the brakets [] you include the text that you want included as the figure caption.  Then Within the parentheses () you include the link to where the image can be found; the image is from the web in the above example.  I have also nested the link to Schmidts paper's DOI within the figure text which is why it appears complicated.  Below is a similar example with a different image.  Remember you will be copying this template and, though trial and error, it will begin to make sense... and if it doesnt, we'll help you in class. 
```

Discuss the emerging understandings in stabilization mechanisms. 

![Figure 2. An emerging understanding of stabilization/destabilization processes has replaced humification theory. For full figure text see Schmidt et al. (2011) DOI: [10.1038/nature10386](https://www.nature.com/articles/nature10386)](https://media.springernature.com/m685/nature-assets/nature/journal/v478/n7367/images/nature10386-f3.2.jpg)

List of important publications:

Schmidt et al. (2011) DOI: [10.1038/nature10386](https://www.nature.com/articles/nature10386)

#Rethinking the Conceptual Theory of SOM Sequestration

As paradigms have shifted, its important to get away from humification terminology and embrace mechanistic theories and conceptualizations of soil carbon stabilization.

![Figure 3. The soil continum model (SCM) is an attempt to bring together disparate conceptualizations of soil carbon storage, which is an important step forward in soil science as we try to determine the best measurement and modeling techniques to predict SOC changes. For full figure text see Lehmann & Kleber (2015) DOI: [10.1038/nature16069](https://www.nature.com/articles/nature16069)](https://media.springernature.com/m685/nature-assets/nature/journal/v528/n7580/images/nature16069-f2.jpg)

List of important publications:

Lehmann & Kleber (2015) DOI: [10.1038/nature16069](https://www.nature.com/articles/nature16069)

#Moving forward: bringing together theory, measurement, and modeling

Theory, measurement, and modeling need to be thought of as a single process to close the loop of the scientific method and create usable products to not just inform further science, but also provide farmers, land managers, policy makers, and the general public a tool to understand soil carbon. 

List of important publications:

Triangle paper
